---
title: "EDA"
author: ""
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: pdf_document
---

We load the necessary packages to execute all the queries.

```{r}
library(questionr)
library(dplyr)
library(factoextra)
library(readxl)
#library(summarytools)
library(compareGroups)
citation("compareGroups")
library("FactoMineR")
citation("FactoMineR")
library("factoextra")
citation("factoextra")
library(DataExplorer)
```

#DATABASE FOR ANALYSIS

We load the database.

```{r}
bd <- read_excel("/Volumes/PENMARTI/ARTICLES/ARTICLE CENS/DB_ANALYSIS.xlsx")
head(bd)
```

We rename the variables to make them easier to identify.

```{r}
noms <- c("Gender","Age group","Country categorized","Type","Highest level of education","Years worked in sports statistics in academia","Profile","Currently belong to a sports statistics research group?","Total citations in GS","Total citations in GS since 2021","Total citations in GS acording H-index","Total citations in GS acording H-index since 2021","Number of citations of the article with most citations","Number of sports studied")
colnames(bd) <- noms
attach(bd)
```

We convert the categorical variables to factors so that they can be handled more easily.

```{r}
# First, we transform the database into a data frame
bd <- data.frame(bd)
# Then, we identify the categorical variables and collect them all into a vector.
x <- c()
for(i in 1:ncol(bd)){
  if(is.character(bd[,i])){
    x <- c(x,i)
  }
}
# Finally, we declare them as factors.
for(i in x) {
  bd[,i]<-as.factor(bd[,i])
}
# Finally, we check that they have been transformed into a vector.
str(bd)
```

**Univariate Analysis**

Let's look at the frequencies and key statistical measures.

*Variable Gender*

```{r}
freq(bd$Gender, cum = F, digits = 2)
```

*Variable Age group*

```{r}
freq(bd$`Age.group`, cum = F, digits = 2)
```

*Variable Country categorized* 

```{r}
freq(bd$`Country.categorized`, cum = F, digits = 2)
```

*Variable Type*

```{r}
freq(bd$Type, cum = F, digits = 2)
```

*Variable Highest level of education*

```{r}
freq(bd$`Highest.level.of.education`, cum = F, digits = 2)
```

*Variable years worked in sports statistics in academia*

```{r}
freq(bd$`Years.worked.in.sports.statistics.in.academia`, cum = F, digits = 2)
```

*Variable Profile*

```{r}
freq(bd$Profile, cum = F, digits = 2)
```

*Variable Currently belong to a sports statistic research group?*

```{r}
freq(bd$`Currently.belong.to.a.sports.statistics.research.group.`, cum = F, digits = 2)
```

*Variable Total citations in GS*

```{r}
summary(bd$Total.citations.in.GS)
sd(bd$Total.citations.in.GS)
```

*Variable Total citations in GS since 2021*

```{r}
summary(bd$Total.citations.in.GS.since.2021)
sd(bd$Total.citations.in.GS.since.2021)
```

*Variable Total citations in GS acording H-index*

```{r}
summary(bd$Total.citations.in.GS.acording.H.index)
sd(bd$Total.citations.in.GS.acording.H.index)
```

*Variable Total citations in GS acording H-index since 2021*

```{r}
summary(bd$Total.citations.in.GS.acording.H.index.since.2021)
sd(bd$Total.citations.in.GS.acording.H.index.since.2021)
```

*Variable Number of citations of the article with most citations*

```{r}
summary(bd$Number.of.citations.of.the.article.with.most.citations)
sd(bd$Number.of.citations.of.the.article.with.most.citations)
```

*Variable Number of sports studied*

```{r}
summary(bd$Number.of.sports.studied)
sd(bd$Number.of.sports.studied)
```

**Bivariate Analysis**

We perform a Shapiro-Wilk test to check if the numerical variables follow a normal distribution.

```{r}
shapiro.test(bd$Total.citations.in.GS)
shapiro.test(bd$Total.citations.in.GS.since.2021)
shapiro.test(bd$Total.citations.in.GS.acording.H.index)
shapiro.test(bd$Total.citations.in.GS.acording.H.index.since.2021)
shapiro.test(bd$Number.of.citations.of.the.article.with.most.citations)
shapiro.test(bd$Number.of.sports.studied)
```

Comparison between the response variable and all other variables.

*Variable Gender*

```{r}
# Do not consider Country, Type, and Research group.
compareGroups(Gender~., data = bd, method = 2, chisq.test.perm = F)
```

```{r}
# Only consider Country, Type, and Research group.
compareGroups(Gender~., data = bd, method = 2, chisq.test.perm = T)
```

*Variable Age group*

```{r}
# Do not consider Country, Type, and Research group.
compareGroups(Age.group~., data = bd, method = 2, chisq.test.perm = F)
```

```{r}
# Only consider Country, Type, and Research group.
compareGroups(Age.group~., data = bd, method = 2, chisq.test.perm = T)
```

*Variable Country categorized*

```{r}
# Do not consider Country, Type, and Research group.
compareGroups(Country.categorized~., data = bd, method = 2, chisq.test.perm = F)
```

*Variable Type*

```{r}
# Do not consider Country, Type, and Research group.
compareGroups(Type~., data = bd, method = 2, chisq.test.perm = F)
```

```{r}
# Only consider Country, Type, and Research group.
compareGroups(Type~., data = bd, method = 2, chisq.test.perm = T)
```

*Variable Highest level of education*

```{r}
# Do not consider Country, Type, and Research group.
compareGroups(Highest.level.of.education~., data = bd, method = 2, chisq.test.perm = F)
```

```{r}
# Only consider Country, Type, and Research group.
compareGroups(Highest.level.of.education~., data = bd, method = 2, chisq.test.perm = T)
```

*Variable Years worked in sports statistics in academia*

```{r}
# Do not consider Country, Type, and Research group.
compareGroups(Years.worked.in.sports.statistics.in.academia~., data = bd, method = 2, chisq.test.perm = F)
```

```{r}
# Only consider Country, Type, and Research group.
compareGroups(Years.worked.in.sports.statistics.in.academia~., data = bd, method = 2, chisq.test.perm = T)
```

*Variable Years worked in sports statistics in academia*

```{r}
# Do not consider Country, Type, and Research group.
compareGroups(Profile~., data = bd, method = 2, chisq.test.perm = F)
```

```{r}
# Only consider Country, Type, and Research group.
compareGroups(Profile~., data = bd, method = 2, chisq.test.perm = T)
```

*Variable Currently belong to a sports statistics research group?*

```{r}
# Do not consider Country, Type, and Research group.
compareGroups(Currently.belong.to.a.sports.statistics.research.group.~., data = bd, method = 2, chisq.test.perm = F)
```

```{r}
# Only consider Country, Type, and Research group.
compareGroups(Currently.belong.to.a.sports.statistics.research.group.~., data = bd, method = 2, chisq.test.perm = T)
```

Comparison between the categorized country and the years worked.

```{r}
createTable(compareGroups(Country.categorized~Years.worked.in.sports.statistics.in.academia, data = bd, chisq.test.perm = F))
createTable(compareGroups(Country.categorized~Years.worked.in.sports.statistics.in.academia, data = bd, chisq.test.perm = F), show.p.mul=TRUE)
```

Comparison between the education level and the years worked.

```{r}
createTable(compareGroups(Years.worked.in.sports.statistics.in.academia~Highest.level.of.education, data = bd, chisq.test.perm = F))
createTable(compareGroups(Years.worked.in.sports.statistics.in.academia~Highest.level.of.education, data = bd, chisq.test.perm = F), show.p.mul=TRUE)
```

*Correlation analysis of the numerical variables*

```{r}
var_num <- bd[9:14]
noms <- c("Total citations","Total citations since 2018","Total citations H-index","Total citations H-index since 2018","Article with most citations", "Number of sports studied")
colnames(var_num) <- noms
```

We create the correlation matrix.

```{r}
noms <- c("Var1","Var2","Var3","Var4","Var5","Var6")
colnames(var_num) <- noms
plot_correlation(var_num, title = "Correlation matrix among the numerical variables")
```

We perform a Principal Component Analysis (PCA).

We visualize the inertia of each component graphically.

```{r}
res.pca <- PCA(var_num, graph = FALSE)
fviz_eig(res.pca, addlabels = TRUE, main = "Graph of variances by each component", xlab = "Components", ylab = "Percentage of explained variance")
```

Now let's look at the exact values of each component.

```{r}
res.pca$eig
```

And finally, we create the PCA plot.

```{r}
fviz_pca_var(res.pca, col.var = "cos2",
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"), 
             repel = TRUE, xlab = "Component 1 (68.83%)", ylab = "Component 2 (15,14%)")
```

Now we remove the three outlier observations and we perform a Principal Component Analysis (PCA) again.

```{r}
bd <- bd[-28,]
bd <- bd[-34,]
bd <- bd[-69,]
```

```{r}
res.pca <- PCA(var_num, graph = FALSE)
fviz_eig(res.pca, addlabels = TRUE, main = "Graph of variances by each component", xlab = "Components", ylab = "Percentage of explained variance")
```

```{r}
res.pca$eig
```